---
title: "Full example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{full-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 10,
  fig.height = 6,
  out.width = "100%",
  dpi = 300,
  comment = "#>"
)
```

```{r, warning = F,  message = FALSE, results = FALSE }
library("tna")
library("tibble")
library("dplyr")
library("gt")
library("seqHMM")
library("TraMineR")
```

Import the data
```{r}
data("engagement", package = "tna")
```

Build the model
```{r}
tna_model <- build_tna(engagement)
```

Plot the model
```{r}
plot(tna_model)
```

Initial probabilities
```{r}
data.frame(`Initial prob.` = tna_model$inits[[1]]) |> 
  rownames_to_column("Engagement state") |> arrange(desc(`Initial.prob.`)) |> 
  gt() |> fmt_percent()
```

Transition probabilities
```{r}
tna_model$transits[[1]] |> data.frame() |> 
  rownames_to_column("From\\To") |> 
  gt() |> fmt_percent()
```

Centrality measures
```{r, fig.width=9,fig.height=6}
centrality_measures = c("Betweenness", "Closeness", "InStrength", "OutStrength")

cents_withoutloops <- centralities(tna_model, measures = centrality_measures, 
                                   loops = FALSE, normalize = TRUE)

plot(cents_withoutloops, ncol = 2, reorder = T,
     point_color = tna_model$colors, line_color = tna_model$colors)  
```


# Clustering

Initialization
```{r, warning = F, cache = TRUE}
set.seed(265)

Nvar <- length(tna_model$labels)
Nclusters <- 3
trans_probs <- simulate_transition_probs(Nvar, Nclusters)
init_probs <-list(c(0.70, 0.20, 0.10), 
                  c(0.15, 0.70, 0.15), 
                  c(0.10, 0.20, 0.70))
```

Building and fitting the model (this might take a while...)
```{r, warning = F, cache = TRUE}
set.seed(265)
modelTrans <- build_mmm(engagement, transition_probs = trans_probs,
 initial_probs = init_probs)

controlEM = list(restart = list(times = 100, n_optimum = 101))

fitModelTrans <- fit_model(
 modelTrans, 
 global_step = TRUE,
 control_global = list(algorithm = "NLOPT_GD_STOGO_RAND"),
 local_step = TRUE,
 threads = 60,
 control_em = controlEM)
```

Plotting each cluster sequence
```{r, fig.width=9, fig.height=5}
clusternames <- c("Cluster 1", "Cluster 2", "Cluster 3")

cluster_names(fitModelTrans$model) <- clusternames
sumamaryx <- summary(fitModelTrans$model)
seqdplot(engagement, group = sumamaryx$most_probable_cluster, rows = 2, cols = 2)  
seqIplot(engagement, group = sumamaryx$most_probable_cluster, rows = 2, cols = 2)   
```



Creating a new model with the cluster information
```{r}
tna_model_clus <- build_tna(fitModelTrans$model)
```


Initial probabilities
```{r}
data.frame(tna_model_clus$inits) |> 
  rownames_to_column("Engagement state") |>  
  gt() |> fmt_percent()
```

Transition probabilities

```{r}
tna_model_clus$transits[[1]] |> data.frame() |> 
  rownames_to_column("From\\To") |> gt() |>
  tab_header(title = names(tna_model_clus$transits)[1]) |>
  fmt_percent()
tna_model_clus$transits[[2]] |> data.frame() |> 
  rownames_to_column("From\\To") |> gt() |>
  tab_header(title = names(tna_model_clus$transits)[2]) |>
  fmt_percent()
tna_model_clus$transits[[3]] |> data.frame() |> 
  rownames_to_column("From\\To") |> gt() |>
  tab_header(title = names(tna_model_clus$transits)[3]) |>
  fmt_percent()
```

 

Plotting the cluster transitions
```{r, fig.width=10, fig.height=3.2}
layout(t(1:3))

plot(tna_model_clus, cluster = 1, edge.label.cex = 2, vsize = 18, 
     title = clusternames[1], title.cex = 1.15)
plot(tna_model_clus, cluster = 2, edge.label.cex = 2, vsize = 18, 
     title = clusternames[2], title.cex = 1.15)
plot(tna_model_clus, cluster = 3, edge.label.cex = 2, vsize = 18, 
     title = clusternames[3], title.cex = 1.15)
```


Cluster centralities
```{r, fig.width=9, fig.height=4}
layout(t(1:2))
centralities_per_cluster <- centralities(tna_model_clus, normalize = T, 
                                         measures = centrality_measures) 
plot(centralities_per_cluster, scales = "free")
```
